---
title: "Lab Week 8"
author: "Sandy Carter"
date: "3/8/2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# 1. Point Pattern Analysis Oil Spill Texas

### Load packages
```{r, message = FALSE}
library(tidyverse)
library(sf)
library(tmap)
library(leaflet)
library(spatstat)
library(maptools)
```

### Make column graph of Texas Oil Spills
```{r, message=FALSE}
oil_spills <- read_csv("oil_spills.csv")

df <- oil_spills %>% 
  filter(`Accident State` == "TX" & `Accident Year` < 2017) %>% 
  group_by(`Accident Year`) %>%  # Create groups by different years so we can find net oil loss by year
  summarise(Loss = sum(`Net Loss (Barrels)`))

colnames(df) <- c("Year", "Loss")

ggplot(df, aes(x = Year, y = Loss)) +
  geom_col()
# Loss is in barrels. This is about 1/2 as much as Exxon Valdez. (but land to marine is apples/oranges)
```

### Leaflet plot of spill locations in TX in 2016

```{r}
df_loc <- oil_spills %>% 
  filter(`Accident State` == "TX" & `Accident Year` == 2016) %>% 
  select(Latitude, Longitude, `Net Loss (Barrels)`)

colnames(df_loc) <- c("latitude", "longitude", "net_loss")

oil_sf <- st_as_sf(df_loc, coords = c("longitude", "latitude"), crs = 4326)   # Change lat and long into spatial data for R; makes spatial information sticky 

leaflet(oil_sf) %>% 
  addTiles() %>% 
  addMarkers()

```

### Make tmap of oil spills with Texas State shape file (to prep for spatial point pattern analysis)
```{r}
states <- st_read(dsn = ".", layer = "states")

tex_border <- states %>% 
  filter(STATE_NAME == "Texas") %>% # select texas
  st_transform(4326) # Transform coordinate reference system (need polygons to align correctly); using transform b/c it already had one--not setting it, but changing it

#plot(tex_border)

tm_shape(tex_border) +
  tm_polygons() +
  tm_shape(oil_sf) +
  tm_dots(size = 0.3)
```
  
**Are these spatially random or not? Does this point pattern follow CRS? (Doing this on oil spills isn't the best in that these are dependent on oil locations, but we're going to use it as an example here)**

### Convert the data to spatial points pattern (combination of point data and the bounding window)
```{r}
# identify points
spill_sp <- as(oil_sf, "Spatial") # convert back from simple features to dataframe
spill_ppp <- as(spill_sp, "ppp") #Change class to ppp (point pattern analysis)

# identify bounding window
tx_sp <- as(tex_border, "Spatial")
tx_owin <- as(tx_sp, "owin") #Take texas spatial and set it as window for ppp

# Put points and bounding window together
all_ppp <- ppp(spill_ppp$x, spill_ppp$y, window = tx_owin) #x and y are corresponding lat and long. R omits points that don't align within the polygon

```

### Density plot
```{r}
plot(density(all_ppp, sigma=0.4)) # Caution: density plots are really easy to make, but they can change stories based on sigma values. Sigma = bandwidth 
```

### Quadrat test for spatial evenness
Are oil spills evenly distributed throughout the state?

```{r}
oil_qt <- quadrat.test(all_ppp, nx = 5, ny = 5) # add ppp, set quadrat sizes--the fact you can set these is why it's dangerous. 5 regions horizontally (nx), 5 regions vertically(5)--breaks up state of texas into 5 regions in each way. Counts events in each of those quadrats and tests that with chi-square
oil_qt
# Result: The data are not spatially evenly distributed
    # Small p-value. look at the state, it doesn't really look like CSR (see strong clusterings in some areas). This tests null hypothesis of CSR. Really high chi-square and small p-value = hard reject null hypothesis. retain alternative hypothesis that it's not CSR, not evenly distributed. 

plot(all_ppp)
plot(oil_qt, add=TRUE,cex = 0.4) #upper right = expected; upper left = actual count; bottom = standardized measure of how different expected and actual are
```

### Nearest Neighbor Analysis
If clustered, nearest neighbor will be closer than under CSR.  
Two ways to evaluate:  
(1) Consider single closest nearest neighbor, calculate g fcn (proportion of point pairs within distance r).    
(2) Consider how close points are overall, calculate L and K functions (proportion of point density within bubble around each point as bubble gets larger).  

#### G function for nearest neighbor analysis

```{r}
r <- seq(0,1, by = 0.01) #Make sequence of values for r (this will be a different set for each dataset--need a span that considers span of distance that covers all of them)

oil_gfun <- envelope(all_ppp, fun = Gest, r = r, nsim = 100) # Simulate CSR data to compare the data to, using the r window and # points. Use a poisson distribution. 
#outcome df: r, observation for each of those ranges (# point pairs--actual data), theoretical = theoretical values based on CSR, low and high confidence intervals) 

ggplot(oil_gfun, aes(x=r, y = obs)) +
  geom_line(color = "black") +
  geom_line(aes(x=r, y = theo), color = "red")
# Shows us that our observed data has a higher proportion of point pairs with nearest neighbors at shorter distances compared to CSR data (i.e., more clustering)
```

#### L function for nearest neighbor analysis (standardized version of Ripley's K function)
```{r}
r2 <- seq(0,3, by = 0.5) #chose large increment to avoid having the simulation take 10 min to run

oil_lfun <- envelope(all_ppp, fun = Lest, r = r2, nsim=20, global = TRUE)

ggplot(oil_lfun, aes(x = r2, y = obs)) +
  geom_line(color = "black") +
  geom_line(aes(x = r2, y = theo), color = "blue")
# takeaway similar as G: higher densities of neighbors that exist at small distances than CSR
```
Recap:
look at map (doesn't seem CSR); look at quadrat map (not even); look at g and l fcn (greater clustering)


